---
title: "EchoviewR - Basic Functionality"
author: "Sven Gastauer"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Introduction  


The R package EchoviewR - a free interface between Echoview (R) and R using COM scripting.

### Copyright and Licence  

<pre><code>
Copyright (C) 2015 Lisa-Marie Harrison, Martin Cox, Georg Skaret and Rob Harcourt.

This file is part of EchoviewR.

EchoviewR is free software: you can redistribute it and/or modify
it under the terms of the GNU General Public License as published by
the Free Software Foundation, either version 3 of the License, or
(at your option) any later version.

EchoviewR is distributed in the hope that it will be useful,
but WITHOUT ANY WARRANTY; without even the implied warranty of
MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
GNU General Public License for more details.

You should have received a copy of the GNU General Public License
along with EchoviewR.  If not, see <http://www.gnu.org/licenses/>.</pre></code>

</pre></code>

This package is open for community development and we encourage users to extend the package as they need. We are not liable for any losses when using EchoviewR.

If using EchoviewR, please cite as:

```{r eval=FALSE}
citation('EchoviewR')
```

## EchoviewR - Installation guide

***

*Installing all dependencies:*  

EchoviewR currently has a few dependencies:

* [fields](https://cran.r-project.org/web/packages/fields/index.html) 
* [sp](https://cran.r-project.org/web/packages/sp/index.html) 
* [geosphere](https://cran.r-project.org/web/packages/geosphere/index.html) 
* [maptools](https://cran.r-project.org/web/packages/maptools/index.html) 
* [RDOCMClient](http://www.omegahat.net/RDCOMClient/) 
* [rgeos](https://cran.r-project.org/web/packages/rgeos/index.html) 
* [lubridate](https://cran.r-project.org/web/packages/lubridate/vignettes/lubridate.html)

```{r eval=FALSE}
install.packages(c('fields', 'sp', 'geosphere', 'maptools', 'RDCOMClient', 'rgeos', 'lubridate'))
```

Sometimes the installation of [RDCOMClient](https://github.com/omegahat/RDCOMClient) from CRAN fails. If this is the case, run:

```{r eval=FALSE}
install.packages("RDCOMClient", repos = "http://www.omegahat.net/R")
```

If you intend to create variables from within R you will need to install RDCOMMServer as well:  

```{r eval=False}
install.packages("RDCOMServer", repos = "http://www.omegahat.net/R")

```
***

*Installing EchoviewR from [Github](https://github.com/AustralianAntarcticDivision/EchoviewR):*  

```{r eval=FALSE}

devtools::install_github('AustralianAntarcticDivision/EchoviewR')
```

## Basic functions

EchoviewR makes startin, closing and using Echoview through easy. In this part of the vignettes some of the basic functions will be demonstrated.
We start by starting an Echoview Instance:  
